Properties to calculate:

energy, mean radial displacement, one-body density, two-body density, learning rate, standard deviation, variation, for RBM, plot the matrix W.
Confidence interval 

Notes results(draft)

All for 2 particles in 3 dimensions.

Biggest Problem!! $\alpha$ will NOT stabilise, only increases when interaction is on....  
And method fluctuates too much, energy does not stabilises. Gradient of local energy is to high, solution? 
Problem with energy being both too high AND too low..

\section{Non-interacting}
Used the analytic solution of the local energy
\subsection{Metropolis}
Run with naive metropolis works fine. 
Monte Carlo cycles = $10^5$


\subsection{Importance}
Run with numerical calculation of the quantum force $F$ is ok, number of Monte Carlo cycles = $10^5$

Energy stabilises after few gradient descent runs, 100 something runs. 

Optimal $\alpha$ is also reached

Importance sampling with analytical solution to the quantum force also works fine. 

Monte Carlo cycles = $10^5$

\section{Weak interacting}
Used the analytic AND numerical solution of the local energy

With a = 0.00433, correct total energy = 3.00345

a = 0.433, total energy = 3.3827
\subsection{Metropolis}
With numerical calculation of the local energy:

Not correct energy

Total energy = 2.998 ish and alpha approx 0.5006

With analytical calculation of the local energy:

Total energy = 2.998 ish and alpha approx 0.5006
But the derivative of the energy is high! So something is wrong with the gradient method??

\subsection{Importance}
Monte Carlo cycles = $10^5$
With numerical calculation of the local energy:

- and numerical solution of quantum force

Not correct energy nor alpha

Energy stabilises at to low, 2.998 and alpha is to high, above 0.5006
 
- analytical solution of quantum force
With my own expression the method does not work at all, energy and alpha waaay to high..
Used Håkons expression of the factor i derivative of $u$, times 2. Do NOT know why this worked. Must find that out.. 

Still not correct energy though, but closer to the correct one. Alpha still to high

Energy goes too high, 3.005 and alpha is to high, approx 0.5033

With analytical calculation of the local energy:

- and numerical solution of quantum force

Energy around 3.011, alpha 0.50326, not entirely correct...
 
- analytical solution of quantum force

Energy around 3.01, alpha 0.50326, not entirely correct...

\section{Strong interaction}
\subsection{Metropolis}


\subsection{Importance}


\subsection{Gibbs}