Some introduction containing historical information and stuff, and motivation for this section.

\section{Ideal Bose gas}
Bosons are quantum particles with integer spin, meaning they follow Bose-Einstein statistics. 
Elementary: Higgs boson, photon, gluon, gravition

Composite: hydrogen, mesons, nucleus of deuterium

Consider a dilute Bose gas where the inter-particle distance is much greater than the scattering length. It can be approximated by a mean field theory. We describe it by a simple hard-sphere effective potential. (Master Joachim)

Bogoliubov - 

Ground state energy

\section{The Bose-Einstein Condensation}
The ability of bosons at low temperature to collectively inhabit the same energy state, compared to fermions who are restricted to the Pauli exclusion principle.

Free Bose gas - 

\begin{equation} \label{eq:hamilt_free_bose_gas}
H^{free}_V = \int_V dx \frac{1}{2m} \nabla a*(x). \nabla a(x)
\end{equation}

BEC in interaction - 

BEC in trap - 

The Bose-Einstein condensation ground state can be described by the Gross-Pitaevskii equation, which uses the Hartree-Fock approximation and the pseduopotential intereaction model. (wiki) 
A a many-body boson system, the Gross-Pitaevskii approach is a mean-field approach. (MBBS)
If the spacing between the particles are greater than the scattering length (in the so-called dilute limit), then one can approximate the true interaction potential that features in this equation by a psedopotential. READ more about why this is.

The wave function in the Hartree-Fock approximation of a system of N bosons is represented by the product of single-particle wave functions $\psi$

\begin{equation}
\psi(r_1, r_2, \cdots, r_N) = \psi(r_1) \psi(r_2) \cdots \psi(r_N)
\end{equation}

The pseudopotential model Hamiltonian of the system is given as 

\begin{equation}
H = \sum_{i=1}^N \left( -\frac{\hbar^2}{2m} \frac{\partial^2}{\partial r_i^2} + V(r_i) \right) + \sum_{i<j} \frac{4 \pi \hbar^2 a_s}{m} \delta (r_i - r_j)
\end{equation}

m is the mass of the boson, V is the external potential, $a_s$ is the boson-boson scattering length and $\delta(r)$ the Dirac delta-function. This is an effective potential, the original  with modifications 

The Gross-Pitaevskii equation reads, 

\begin{equation}
\left( -\frac{\hbar^2}{2m} \frac{\partial^2}{\partial r^2} + V(r) + \frac{4 \pi \hbar^2 a_s}{m}|\psi(r)|^2 \right) \psi(r) = \mu \psi (r)
\end{equation}

Se 104 MBBS

Solutions: free particle V(r) = 0

Soliton?

Thomas-Fermi approximation

Bogoliubov approximation

Interacting Bose gas - Landau's phenomenological theory of superfluidity. Based on the idea that a quantum liquid remains a classical fluid even at zero temperature  and that the classical hydrodynamical laws remain valid. (Many body boson system.) 
Many models have been suggested to create a microscopical theory of the superfluidity. Vortex ring model, hard-sphere model, guassian cluster approach. (wiki)

The properties of these liquids are described in terms of the spectrum of the collective excitations.

\section{The wave function}
We are studying a system of two electrons confined in a harmonic oscillator trap described by the Hamiltonian 

\begin{equation}\label{eq:hamilt}
\hat{H} = \sum_{i=1}^N \left( - \frac{1}{2} \nabla_i^2 + \frac{1}{2} \omega^2 r_i^2 \right) + \sum_{i<j} \frac{1}{r_{ij}}
\end{equation}

where the first sum is the standard harmonic oscillator part and the last is the interacting part between the electrons and $N$ represent the number of particles. $\omega$ is the oscillator frequency of the trap and $r_i$ is the position of particle $i$, whereas $r_{ij}$ is the distance between the particles and given as $r_{ij} = |\mathbf{r_i} - \mathbf{r_j}|$. \\
Our wave function is given from the energy of the restricted Boltzmann machine, see \eqref{eq:E_rbm}, which is the joint energy functional between the visible and hidden nodes. From the marginal probability of the joint probability distribution, see \eqref{eq:F_rbm_marg}, we get our wave equation.                     

\begin{align}\label{eq:F_rbm}
\Psi(X) &= F_{rbm}(X) \\
&= \frac{1}{Z} \exp \left( -\sum_{i}^{M} \frac{(X_i - a_i)^2}{2 \sigma^2} \right) \prod_{j}^{N} \left( 1 + \exp \left( b_j + \sum_{i}^{M} \frac{X_i \omega_{ij}}{\sigma^2} \right) \right)
\end{align}

Here $Z$ is the partition function, $X_i$ represents the visible nodes running up to $M$, and $a_i$ and $b_j$ are the biases described in the section below, \eqref{sec:rbm}, where number of hidden nodes $j$ runs up to $N$. 
$\omega_{ij}$ is an $M \times N$ matrix holding the weights connecting the visible nodes with the hidden and $\sigma$ is the standard deviation of the noise in our model. \\

\section{Interaction}